---
title: "Exercise 21"
author: "Diana Solis, Shakiba Fadaei, Todor Gitchev"
date: "5/19/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

We calculate the Bayes classifier for one explanatory variable (feature) under the assumption that the class-conditional probability densities are normal. We want to distinguish between two fish species, sea bass (class 0) and salmon (class 1). Assume the class-conditional probability densities for the length x of the fishes are given by the following normal distributions:

$X | Y=0 \sim \mathcal{N}\left(\mu=7, \sigma^{2}=1\right)$ for sea brass
and
$X | Y=1 \sim \mathcal{N}\left(\mu=3, \sigma^{2}=1\right)$ for salmon

$$
f(x)=\frac{1}{\sqrt{2 \pi \sigma^{2}}} e^{-\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^{2}}
$$

```{r}
length <- seq(0, 12, by=0.1) # lengt measurments

fx_y0 <- function(length) dnorm(length, mean = 7, sd = 1)
fx_y1 <- function(length) dnorm(length, mean = 3, sd = 1)
```

# a)
We have prior class probabilities:
 $\pi_0 = P[Y=0] = \pi_1 = P[Y=1] = 1/2$

posterior class probabilities: $P\left(y_{j} | x\right)=\frac{p\left(x | y_{j}\right) \pi_{j}}{p(x)} \quad j=0,1$

evidence: $p(x)=\sum_{i=0}^{1} p\left(x | y_{i}\right) \pi_{i}$

a posteriory: $p\left(x | y_{i}\right) \pi_{i}$


```{r}
fx_evidence <- function(length, pi_0, pi_1) pi_0*fx_y0(length) + pi_1*fx_y1(length)

plot_class_prob <- function(pi_0, pi_1) {
  plot(length, fx_y0(length), main = "Class-conditional probability",  ylab="prob. density", ylim=c(0,0.5), type="l", lty = 1, col="cyan3")
  lines(length, pi_0*fx_y0(length), type="l", lty = 2, col="cyan3")
  lines(length, fx_y1(length),type="l", lty = 1, col="brown1")
  lines(length, pi_1*fx_y1(length), type="l", lty = 2, col="brown1")
  lines(length, fx_evidence(length, pi_0, pi_1), type="l", lty = 1, col="darkblue")
  decision_boundary <- (log(pi_1/pi_0) + 40) / 8
  abline(v=decision_boundary, lty=2, col="black" )
  text(decision_boundary, 0.5, "decision boundary", cex= 0.7)

  legend("topright", legend = c("sea bass", "salmon", "a posteriori", "evidence"), col=c("cyan3", "brown1", "gray", "darkblue"), lty = c(1, 1, 2, 1), cex = 0.7)
} 

plot_class_prob(0.5, 0.5)
```

# b)
We have prior class probabilities:
 $\pi_0 =P [Y = 0] = 2/3$, $\pi_1 =P [Y = 1] = 1/3$
 

```{r}
plot_class_prob(2/3, 1/3)
```

# c)
 Decision boundary:
$$
\pi_0 \cdot \frac{1}{\sqrt{2 \pi}} e^{-\frac{(x-7)^2}{2}} = \pi_1 \cdot \frac{1}{\sqrt{2 \pi }} e^{-\frac{(x-3)^2}{2}}
$$

$$
\pi_0 \cdot e^{-(x-7)^2}= \pi_1 \cdot e^{-(x-3)^2}
$$

$$
 log(\pi_0) - {\left(x-7\right)^{2}} = log(\pi_1) - {(x-3)^2}
$$

$$
 log(\pi_0)  - x^2 +14x -49 = log(\pi_1) - x^2 + 6x - 9 => x=\frac{log(\pi_1/\pi_0) + 40}{8}
$$

